---
title: "Introduction to R and Rstudio"
subtitle: "Session  - objects"
execute:
    echo: true
    eval: true
---

## Creating an object

```{r}
#| label: "libs"
#| include: false
library(readr)
library(dplyr)
```

```{r }
#| label: "load-data"
#| echo: false
beds_data <- read_csv(url("https://raw.githubusercontent.com/nhs-r-community/intro_r_data/main/beds_data.csv"), 
                      col_types = cols(date = col_date(format = "%d/%m/%Y")), 
                      skip = 3)
```

Objects that are created from imported data appear in the Environment pane  

It is possible to also create objects from code


## Temporary and temporary

The following code exists only as the code is run and the underlying object `beds_data` is never changed

```{r}

beds_data |> 
  summarise(total_beds = sum(beds_av, na.rm = TRUE),
            total_occupancy = sum(occ_av, na.rm = TRUE),
            .by = org_name) |> 
  mutate(perc_occ = total_occupancy / total_beds) |> 
  arrange(desc(perc_occ))

```

## Saving as an object

In R the assign operator `<-` is conventionally used instead of `=` although that will work

Shortcut key is `Alt and -`

```{r}

bed_occupancy <- beds_data |> 
  summarise(total_beds = sum(beds_av, na.rm = TRUE),
            total_occupancy = sum(occ_av, na.rm = TRUE),
            .by = org_name) |> 
  mutate(perc_occ = total_occupancy / total_beds) |> 
  arrange(desc(perc_occ))

```

## Naming style

The way names are written out is a question of style but it's best to be consistent. 

Other ways of writing names will work but are best avoided like `ALLCAPS` and `With Spaces`

```{r}
#| eval: false
camelCase # first letter is small case

PascalCase # every letter is capital

snake_case # lower case and words are separated with underline

kebab-case # lower case and hyphen, used in RMarkdown but not R scripts
```

## End session