---
title: "Introduction to R and Rstudio"
title-slide-attributes: 
  data-background-color: "#43464B"
subtitle: "Session  - styling code"
author: "Zoë Turner"
format:
  revealjs:
    theme: [default, nhsr-quarto.scss]
    logo: https://raw.githubusercontent.com/nhs-r-community/assets/main/logo/nhsr-logo.svg
    code-link: true
    css: nhsr-quarto.scss
    self-contained: true
    preview-links: true
execute:
  echo: true
  eval: false
---

## Code style

```{r}
#| echo: false
#| label: "libs"
#| include: false
library(readr)
library(countdown)

```

```{r}
#| echo: false 
beds_data <- read_csv("https://raw.githubusercontent.com/nhs-r-community/intro_r_data/main/beds_data.csv", 
    col_types = cols(date = col_date(format = "%d/%m/%Y")), 
    skip = 3)
```

An updated version of R4DS now includes a section on [code style](https://r4ds.hadley.nz/workflow-style.html) 

It particularly mentions the package {styler} which, when run becomes an addin to RStudio, and automatically applies the [tidyverse style](https://style.tidyverse.org/) to code

```{r}
install.packages("styler")
```

## Have a go - look at automatic indentation

RStudio also automatically indents according to the function bracket positions

- Type `beds_data` then press return, where does the cursor go?
- Type `beds_data |> ` then return, where does the cursor go now?
- Type `beds_data |> select(org_name, org_code)` and put the select on a new line, then put `org_code` onto a new line.
- Copy the following to the Editor:

```{r}
beds_data |> 
select(org_code,
everything)
```

highlight the three lines and use the keyboard shortcut `Ctrl+i`, what happens?

```{r}
#| echo: false
#| eval: true
countdown::countdown(minutes = 8,
                     color_border = "#005EB8",
                     color_text = "#005EB8",
                     color_running_text = "white",
                     color_running_background = "#005EB8",
                     color_finished_text = "#005EB8",
                     color_finished_background = "white",
                     margin = "0.9em",
                     font_size = "2em")
```

## Space code

Using spaces helps with readability

The code below will run as R doesn't need spaces to work:

```{r}
beds_data|>group_by(org_name)|>summarise(total_beds=sum(beds_av,na.rm=TRUE),total_occupancy=sum(occ_av,na.rm=TRUE))|>mutate(perc_occ=total_occupancy/total_beds)|>arrange(desc(perc_occ))
```

Putting the spaces in can be effort after it's been written

## Have a go - {styler}

Firstly copy the code without spaces to the document

```{r}
beds_data|>group_by(org_name)|>summarise(total_beds=sum(beds_av,na.rm=TRUE),total_occupancy=sum(occ_av,na.rm=TRUE))|>mutate(perc_occ=total_occupancy/total_beds)|>arrange(desc(perc_occ))
```

With {styler} loaded `library(styler)` go to the Add in drop down menu and select `Style Active File`

```{r}
#| echo: false
#| eval: true
countdown::countdown(minutes = 7,
                     color_border = "#005EB8",
                     color_text = "#005EB8",
                     color_running_text = "white",
                     color_running_background = "#005EB8",
                     color_finished_text = "#005EB8",
                     color_finished_background = "white",
                     margin = "0.9em",
                     font_size = "2em")
```

## End session