---
title: "Introduction to R and Rstudio"
subtitle: "Session - {janitor} clean data"
execute:
  echo: true
  eval: false
---


## Specific packages to clean data

Packages like {janitor} have functions to do a lot of the cleaning required for data like:

:::incremental
- Remove blank rows and columns
- Change Excel serial dates to read dates
- Standardise column names and remove spaces
:::

## Example of cleaning column headers

```{r}
library(readr)
by_ethnicity <- read_csv("https://www.ethnicity-facts-figures.service.gov.uk/culture-and-community/digital/internet-use/latest/downloads/by-ethnicity.csv")
```

## Changing the column names

Removes spaces and changes the `%` to a word

```{r}
library(janitor)

by_ethnicity |> 
  clean_names()
```

## Example of removing blank rows and columns

```{r}
# Add in blank row and column

library(dplyr)
library(janitor)
by_ethnicity_blank <- by_ethnicity |> 
  mutate(blank_column = NA) |> # Blank column
  add_row() # Blank row

by_ethnicity_blank |> 
  remove_empty(which = c("rows", "cols"))
```

## End session